---
title: "STATS 102B HW2"
author: "Joshua Susanto"
date: "2023-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(tidyverse)
```

## Problem 1: Recall Problem 1 in Homework 1.

Consider the correlation coefficient r between two random variables X and Y. Recall that it is a measure of linear association between X and Y and takes values in the interval [−1, 1]; i.e., r(X,Y) ∈ [−1, 1].

### (a) Use the code from Homework 1 to simulate data from a bivariate normal distribution with mean vector µ = [0 0] and correlation matrix

```{r}
R <- matrix(c(1,'r','r',1),2,2)
R
```

for the following cases:

1. Sample size n ∈ {50, 200} and correlation coefficient r = 0

```{r}
set.seed(405568250)
n <- c(50, 200)
mu <- c(0, 0)
sigma <- matrix(c(1, 0, 0, 1), 2, 2)
par(mfrow = c(1,2))

for(i in n){
  plot(mvrnorm(i, mu, sigma), pch = 16, cex.main = 0.85,
         xlab = "X_1", ylab = "X_2", main = paste(i, "Bivariate Normal Draws: r = 0"))
  points(x = mu[1], y = mu[2], col = "orange", pch = 4, cex = 2)
}
```


2. Sample size n ∈ {50, 200} and correlation coefficient r = 0.5

```{r}
set.seed(405568250)
n <- c(50, 200)
mu <- c(0, 0)
sigma <- matrix(c(1, 0.5, 0.5, 1), 2, 2)
par(mfrow = c(1,2))

for(i in n){
  plot(mvrnorm(i, mu, sigma), pch = 16, cex.main = 0.85,
         xlab = "X_1", ylab = "X_2", main = paste(i, "Bivariate Normal Draws: r = 0.5"))
  points(x = mu[1], y = mu[2], col = "orange", pch = 4, cex = 2)
}
```

3. Sample size n ∈ {50, 200} and correlation coefficient r = 0.85

```{r}
set.seed(405568250)
n <- c(50, 200)
mu <- c(0, 0)
sigma <- matrix(c(1, 0.85, 0.85, 1), 2, 2)
par(mfrow = c(1,2))

for(i in n){
  plot(mvrnorm(i, mu, sigma), pch = 16, cex.main = 0.85,
         xlab = "X_1", ylab = "X_2", main = paste(i, "Bivariate Normal Draws: r = 0.85"))
  points(x = mu[1], y = mu[2], col = "orange", pch = 4, cex = 2)
}
```


### (b) Obtain the following Bootstrap Confidence Intervals for the correlation coefficient r for the three cases above. 

```{r}
bootsampling <- function(x, boot.replicates = B) {
  x = as.matrix(x)
  nx = nrow(x)
  bootsamples = replicate(boot.replicates, x[sample.int(nx, replace = TRUE), ])
}
```


Calculate the length and the shape of each type of Bootstrap CI and report them
as well.

Discuss how you selected the number of bootstrap replicates B.

Comment on the results; in particular how the various bootstrap CI behave as a function of the sample size n, and the value of the correlation coefficient r.




In general, we will give the shape of a confidence interval by measuring the symmetry of the CI with respect to the sample statistic. This will be done by finding the ratio of the distance between the UB/LBs and our statistic.

$$Ratio_{sym} = \frac{|UB - \hat\theta|}{|LB - \hat\theta|}$$


**1. Normal Bootstrap CI**

```{r}
set.seed(405568250)

n = c(50, 200) # Size of a single sample
B = c(200, 500, 1000, 2000, 5000, 10000) # Number of bootstrap replicates
r = c(0,0.5,0.85)

normal_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
zval = qnorm(p = alpha/2, lower.tail = FALSE)

for (i in r) {
  for (j in n) {
    mu <- c(0, 0)
    sigma <- matrix(c(1, i, i, 1), 2, 2)
    x <- mvrnorm(j, mu, sigma)
    theta_hat <- cor(x)[2]
    for (k in B) {
      boot.samples = bootsampling(x, k)
      xbar = apply(boot.samples, 3, cor)[2,]
      se <- sd(xbar)
      normal_CI = rbind(normal_CI, cbind(theta_hat - zval*se, theta_hat + zval*se)) 
    } 
    normal_CI$B <- B
    normal_CI$N <- rep(j, length(B))
    normal_CI$SampleCoefficient <- rep(theta_hat, length(B))
    normal_CI$Length <- normal_CI$V2 - normal_CI$V1
    normal_CI$Shape <- abs(normal_CI$V2 - theta_hat)/abs(normal_CI$V1 - theta_hat)
    normal_CI <- normal_CI %>% select(SampleCoefficient, N, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
    print(normal_CI)
    normal_CI = data.frame(matrix(ncol = 6, nrow = 0))
  }
}
```

According to our CIs we can see that all of our normal CIs have a symmetric shape. When comparing our CIs to our sample statistic we can that a larger N makes our interval more accurate by lowering the length while still containing the true population coefficient. We don't see much of an impact between different numbers of B in the Normal CI scenario, as they all remain relatively similar sized between different numbers of B. When r changed we see that our CI is centered at a different Sample Coefficient, each one being close to the population coefficient r.

**2. Basic Bootstrap CI**

```{r}
set.seed(405568250)

n = c(50, 200) 
B = c(200, 500, 1000, 2000, 5000, 10000) 
r = c(0,0.5,0.85)

basic_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05

for (i in r) {
  for (j in n) {
    mu <- c(0, 0)
    sigma <- matrix(c(1, i, i, 1), 2, 2)
    x <- mvrnorm(j, mu, sigma)
    theta_hat <- cor(x)[2]
    for (k in B) {
      boot.samples = bootsampling(x, k)
      xbar = apply(boot.samples, 3, cor)[2,]
      basic_CI = rbind(basic_CI, cbind(2*theta_hat - quantile(xbar, probs = (1 - alpha/2)), 2*theta_hat - quantile(xbar, probs = (alpha/2))))
    } 
    basic_CI$B <- B
    basic_CI$N <- rep(j, length(B))
    basic_CI$SampleCoefficient <- rep(theta_hat, length(B))
    basic_CI$Length <- basic_CI$V2 - basic_CI$V1
    basic_CI$Shape <- abs(basic_CI$V2 - theta_hat)/abs(basic_CI$V1 - theta_hat)
    basic_CI <- basic_CI %>% select(SampleCoefficient, N, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
    print.data.frame(basic_CI, row.names = FALSE)
    basic_CI = data.frame(matrix(ncol = 6, nrow = 0))
  }
}
```

When comparing our CIs to our sample statistic we can that a larger N makes our interval more accurate by lowering the length while still containing the true population coefficient. Additionally, an increase in N seems to make our shape more symmetric. We don't see much of an impact between different numbers of B in the Normal CI scenario, as they all remain relatively similar sized between different numbers of B. When r changed we see that our CI is centered at a different Sample Coefficient, each one being close to the population coefficient r.

**3. Percentile Bootstrap CI**

```{r}
set.seed(405568250)

n = c(50, 200) 
B = c(200, 500, 1000, 2000, 5000, 10000) 
r = c(0,0.5,0.85)

percentile_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05

for (i in r) {
  for (j in n) {
    mu <- c(0, 0)
    sigma <- matrix(c(1, i, i, 1), 2, 2)
    x <- mvrnorm(j, mu, sigma)
    for (k in B) {
      boot.samples = bootsampling(x, k)
      xbar = apply(boot.samples, 3, cor)[2,]
      percentile_CI = rbind(percentile_CI, cbind(quantile(xbar, probs = alpha/2), quantile(xbar, probs = (1 - (alpha/2)))))
    } 
    percentile_CI$B <- B
    percentile_CI$N <- rep(j, length(B))
    percentile_CI$SampleCoefficient <- rep(cor(x)[2], length(B))
    percentile_CI$Length <- percentile_CI$V2 - percentile_CI$V1
    percentile_CI$Shape <- abs(percentile_CI$V2 - theta_hat)/abs(percentile_CI$V1 - theta_hat)
    percentile_CI <- percentile_CI %>% select(SampleCoefficient, N, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
    print.data.frame(percentile_CI, row.names = FALSE)
    percentile_CI = data.frame(matrix(ncol = 6, nrow = 0))
  }
}
```


When comparing our CIs to our sample statistic we can that a larger N makes our interval more accurate by lowering the length while still containing the true population coefficient. Additionally, an increase in N seems to make our shape more symmetric. We don't see much of an impact between different numbers of B in the Normal CI scenario, as they all remain relatively similar sized between different numbers of B. When r changed we see that our CI is centered at a different Sample Coefficient, each one being close to the population coefficient r. We also see that a high r may attribute to improving the shape of our CIs while also lowering our length.


**4. Bootstrap-t (Studentized) Bootstrap CI**


```{r}
set.seed(405568250)

n = c(50, 200) 
B = c(200, 500, 1000, 2500) 
r = c(0,0.5,0.85)

studentized_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05

for (i in r) {
  for (j in n) {
    mu <- c(0, 0)
    sigma <- matrix(c(1, i, i, 1), 2, 2)
    x <- mvrnorm(j, mu, sigma)
    theta_hat <- cor(x)[2]
    for (k in B) {
      boot.samples <- bootsampling(x, k)
      theta_b = apply(boot.samples, 3, cor)[2,]
      boot.samples.SE = rep(0, dim(boot.samples)[3])
      se <- sd(theta_b)
      for (l in 1:dim(boot.samples)[3]) {
        iterated.samples <- bootsampling(boot.samples[,,l], 100)
        theta_l <- apply(iterated.samples, 3, cor)[2,]
        boot.samples.SE[l] <- sd(theta_l)
      }
      t_b <- (theta_b - theta_hat)/boot.samples.SE
      qval = quantile(t_b, probs = c(alpha/2, 1 - alpha/2)) 
      studentized_CI = rbind(studentized_CI, cbind(theta_hat - qval[2]*se, theta_hat - qval[1]*se))
    } 
    studentized_CI$B <- B
    studentized_CI$N <- rep(j, length(B))
    studentized_CI$SampleCoefficient <- rep(theta_hat, length(B))
    studentized_CI$Length <- studentized_CI$V2 - studentized_CI$V1
    studentized_CI$Shape <- abs(studentized_CI$V2 - theta_hat)/abs(studentized_CI$V1 - theta_hat)
    studentized_CI <- studentized_CI %>% select(SampleCoefficient, N, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
    print.data.frame(studentized_CI, row.names =  FALSE)
    studentized_CI = data.frame(matrix(ncol = 4, nrow = 0))
  }
}
```


When comparing our CIs to our sample statistic we can that a larger N makes our interval more accurate by lowering the length while still containing the true population coefficient. Additionally, an increase in N seems to make our shape more symmetric. We don't see much of an impact between different numbers of B in the Normal CI scenario, as they all remain relatively similar sized between different numbers of B. When r changed we see that our CI is centered at a different Sample Coefficient, each one being close to the population coefficient r. For out studentized CI, we don't see many effects of r on our shape yet we do still see r lowering length and improving accuracy as it increases.


## Problem 2:

The data set “cats” can be obtain through the following R code:

```{r}
library(MASS)
library(tidyverse)
data(cats)
summary(cats)
```

It contains the body weight (Bwt) in kilograms and the (Hwt) in grams of 47 female and 97 male cats.

### Part (a): Construct the following bootstrap CI for the difference of the body weight means
between female and male cats.

Calculate the length and the shape of each type of Bootstrap CI and report them as well.

Discuss how you selected the number of bootstrap replicates B and comment on the results

**1. Normal Bootstrap CI**

```{r}
set.seed(405568250)
female_weight <- cats$Bwt[cats$Sex == "F"]
male_weight <- cats$Bwt[cats$Sex == "M"]

mean_diff_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05
zval = qnorm(p = alpha/2, lower.tail = FALSE)
theta_hat <- mean(male_weight) - mean(female_weight)

B <- c(200, 1000, 5000, 10000)

for (i in B) {
  boot_samples_male <- bootsampling(male_weight, i)
  boot_samples_female <- bootsampling(female_weight, i)
  mean_diff <- apply(boot_samples_male, 2, mean) - apply(boot_samples_female, 2, mean)
  se <- sd(mean_diff)
  mean_diff_CI = rbind(mean_diff_CI, cbind(theta_hat - zval*se, theta_hat + zval*se)) 
}
mean_diff_CI$B <- B
mean_diff_CI$SampleStatistic <- rep(theta_hat, length(B))
mean_diff_CI$Length <- mean_diff_CI$V2 - mean_diff_CI$V1
mean_diff_CI$Shape <- abs(mean_diff_CI$V2 - theta_hat)/abs(mean_diff_CI$V1 - theta_hat)
mean_diff_CI <- mean_diff_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
mean_diff_CI
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


**2. Basic Bootstrap CI**


```{r}
set.seed(405568250)

mean_diff_basic_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05
theta_hat <- mean(male_weight) - mean(female_weight)


B <- c(200, 1000, 5000, 10000)

for (i in B) {
  boot_samples_male <- bootsampling(male_weight, i)
  boot_samples_female <- bootsampling(female_weight, i)
  mean_diff <- apply(boot_samples_male, 2, mean) - apply(boot_samples_female, 2, mean)
  se <- sd(mean_diff)
  mean_diff_basic_CI = rbind(mean_diff_basic_CI, cbind(2*theta_hat - quantile(mean_diff, probs = (1 - alpha/2)), 2*theta_hat - quantile(mean_diff, probs = (alpha/2))))
}
mean_diff_basic_CI$B <- B
mean_diff_basic_CI$SampleStatistic <- rep(theta_hat, length(B))
mean_diff_basic_CI$Length <- mean_diff_basic_CI$V2 - mean_diff_basic_CI$V1
mean_diff_basic_CI$Shape <- abs(mean_diff_basic_CI$V2 - theta_hat)/abs(mean_diff_basic_CI$V1 - theta_hat)
mean_diff_basic_CI <- mean_diff_basic_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(mean_diff_basic_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shapes are relatively symmetric. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


**3. Percentile Bootstrap CI**


```{r}
set.seed(405568250)

mean_diff_percentile_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05
theta_hat <- mean(male_weight) - mean(female_weight)


B <- c(200, 1000, 5000, 10000)


for (i in B) {
  boot_samples_male <- bootsampling(male_weight, i)
  boot_samples_female <- bootsampling(female_weight, i)
  mean_diff <- apply(boot_samples_male, 2, mean) - apply(boot_samples_female, 2, mean)
  mean_diff_percentile_CI = rbind(mean_diff_percentile_CI, cbind(quantile(mean_diff, probs = alpha/2), quantile(mean_diff, probs = (1 - (alpha/2)))))
}
mean_diff_percentile_CI$B <- B
mean_diff_percentile_CI$SampleStatistic <- rep(theta_hat, length(B))
mean_diff_percentile_CI$Length <- mean_diff_percentile_CI$V2 - mean_diff_percentile_CI$V1
mean_diff_percentile_CI$Shape <- abs(mean_diff_percentile_CI$V2 - theta_hat)/abs(mean_diff_percentile_CI$V1 - theta_hat)
mean_diff_percentile_CI <- mean_diff_percentile_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(mean_diff_percentile_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shapes are relatively symmetric. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


**4. Bootstrap-t (Studentized) Bootstrap CI**

```{r}
set.seed(405568250)

mean_diff_t_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05
theta_hat <- mean(male_weight) - mean(female_weight)

B <- c(200, 1000, 2500, 5000)

for (i in B) {
  boot_samples_male <- bootsampling(male_weight, i)
  boot_samples_female <- bootsampling(female_weight, i)
  mean_diff <- apply(boot_samples_male, 2, mean) - apply(boot_samples_female, 2, mean)
  se <- sd(mean_diff)
  boot.samples.SE = rep(0, dim(boot_samples_male)[2])
  for (l in 1:dim(boot_samples_male)[2]) {
        iterated.samples.male <- bootsampling(boot_samples_male[,l], 100)
        iterated.samples.female <- bootsampling(boot_samples_female[,l], 100)
        theta_l <- apply(iterated.samples.male, 2, mean) - apply(iterated.samples.female, 2, mean)
        boot.samples.SE[l] <- sd(theta_l)
  }
  t <- (mean_diff - theta_hat)/boot.samples.SE
  qval = quantile(t, probs = c(alpha/2, 1 - alpha/2)) 
  mean_diff_t_CI = rbind(mean_diff_t_CI, cbind(theta_hat - qval[2]*se, theta_hat - qval[1]*se))
}
mean_diff_t_CI$B <- B
mean_diff_t_CI$SampleStatistic <- rep(theta_hat, length(B))
mean_diff_t_CI$Length <- mean_diff_t_CI$V2 - mean_diff_t_CI$V1
mean_diff_t_CI$Shape <- abs(mean_diff_t_CI$V2 - theta_hat)/abs(mean_diff_t_CI$V1 - theta_hat)
mean_diff_t_CI <- mean_diff_t_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(mean_diff_t_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shapes and length of CI have stayed relatively symmetric. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


### Part (b): Using your code from Problem 1 above to construct the following bootstrap CI for the correlation coefficient between the body weight and the heart weight of female cats.

Calculate their length and their shape and report those as well for each type of bootstrap CI.

Discuss how you selected the number of bootstrap replicates B and comment on the results.

**1. Normal Bootstrap CI**

```{r}
set.seed(405568250)

female_cats <- cats %>% filter(Sex == 'F') %>% select(Bwt, Hwt) %>% data.matrix()
B = c(200, 500, 1000, 2000, 5000, 10000) 
female_normal_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
zval = qnorm(p = alpha/2, lower.tail = FALSE)
theta_hat <- cor(female_cats)[2]

for (k in B) {
  boot.samples.female = bootsampling(female_cats, k)
  sample_stats = apply(boot.samples.female, 3, cor)[2,]
  se <- sd(sample_stats)
  female_normal_CI = rbind(female_normal_CI, cbind(theta_hat - zval*se, theta_hat + zval*se)) 
} 
female_normal_CI$B <- B
female_normal_CI$SampleStatistic <- rep(theta_hat, length(B))
female_normal_CI$Length <- female_normal_CI$V2 - female_normal_CI$V1
female_normal_CI$Shape <- abs(female_normal_CI$V2 - theta_hat)/abs(female_normal_CI$V1 - theta_hat)
female_normal_CI <- female_normal_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(female_normal_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is symmetric as expected.These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


**2. Basic Bootstrap CI**

```{r}
set.seed(405568250)

B = c(200, 500, 1000, 2000, 5000, 10000) 
female_basic_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
theta_hat <- cor(female_cats)[2]

for (k in B) {
  boot.samples.female = bootsampling(female_cats, k)
  sample_stats = apply(boot.samples.female, 3, cor)[2,]
  se <- sd(sample_stats)
  female_basic_CI = rbind(female_basic_CI, cbind(2*theta_hat - quantile(sample_stats, probs = (1 - alpha/2)), 2*theta_hat - quantile(sample_stats, probs = (alpha/2)))) 
} 
female_basic_CI$B <- B
female_basic_CI$SampleStatistic <- rep(theta_hat, length(B))
female_basic_CI$Length <- female_basic_CI$V2 - female_basic_CI$V1
female_basic_CI$Shape <- abs(female_basic_CI$V2 - theta_hat)/abs(female_basic_CI$V1 - theta_hat)
female_basic_CI <- female_basic_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(female_basic_CI, row.names = FALSE)
```

When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is slightly asymmetric with a larger range of values being between our sample statistic and UB. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.

**3. Percentile Bootstrap CI**

```{r}
set.seed(405568250)

B = c(200, 500, 1000, 2000, 5000, 10000) 
female_percentile_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
theta_hat <- cor(female_cats)[2]

for (k in B) {
  boot.samples.female = bootsampling(female_cats, k)
  sample_stats = apply(boot.samples.female, 3, cor)[2,]
  female_percentile_CI = rbind(female_percentile_CI, cbind(quantile(sample_stats, probs = alpha/2), quantile(sample_stats, probs = (1 - (alpha/2))))) 
} 
female_percentile_CI$B <- B
female_percentile_CI$SampleStatistic <- rep(theta_hat, length(B))
female_percentile_CI$Length <- female_percentile_CI$V2 - female_percentile_CI$V1
female_percentile_CI$Shape <- abs(female_percentile_CI$V2 - theta_hat)/abs(female_percentile_CI$V1 - theta_hat)
female_percentile_CI <- female_percentile_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(female_percentile_CI, row.names = FALSE)
```

When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is now slightly asymmetric with a larger range of values being between our sample statistic and LB along with a larger degree of asymmetry. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.

**4. Bootstrap-t (Studentized) Bootstrap CI**

```{r}
set.seed(405568250)

B = c(200, 500, 1000, 2500) 
female_studentized_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05
theta_hat <- cor(female_cats)[2]

for (k in B) {
  boot.samples.female <- bootsampling(female_cats, k)
  theta_b = apply(boot.samples.female, 3, cor)[2,]
  boot.samples.SE = rep(0, dim(boot.samples.female)[3])
  se <- sd(theta_b)
  for (l in 1:dim(boot.samples.female)[3]) {
    iterated.samples <- bootsampling(boot.samples.female[,,l], 100)
    theta_l <- apply(iterated.samples, 3, cor)[2,]
    boot.samples.SE[l] <- sd(theta_l)
  }
  t_b <- (theta_b - theta_hat)/boot.samples.SE
  qval = quantile(t_b, probs = c(alpha/2, 1 - alpha/2)) 
  female_studentized_CI = rbind(female_studentized_CI, cbind(theta_hat - qval[2]*se, theta_hat - qval[1]*se))
} 
female_studentized_CI$B <- B
female_studentized_CI$SampleStatistic <- rep(theta_hat, length(B))
female_studentized_CI$Length <- female_studentized_CI$V2 - female_studentized_CI$V1
female_studentized_CI$Shape <- abs(female_studentized_CI$V2 - theta_hat)/abs(female_studentized_CI$V1 - theta_hat)
female_studentized_CI <- female_studentized_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(female_studentized_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is now slightly asymmetric with a larger range of values being between our sample statistic and LB. We see a slightly larger CI size as well on average. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


### Part (c): Using your code from Problem 1 above to construct the following bootstrap CI for the correlation coefficient between the body weight and the heart weight of male cats.

Calculate their length and their shape and report those as well for each type of
bootstrap CI.

Discuss how you selected the number of bootstrap replicates B and comment on
the results.

**1. Normal Bootstrap CI**

```{r}
set.seed(405568250)

male_cats <- cats %>% filter(Sex == 'M') %>% select(Bwt, Hwt) %>% data.matrix()
B = c(200, 500, 1000, 2000, 5000, 10000) 
male_normal_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
zval = qnorm(p = alpha/2, lower.tail = FALSE)
theta_hat <- cor(male_cats)[2]

for (k in B) {
  boot.samples.male = bootsampling(male_cats, k)
  sample_stats = apply(boot.samples.male, 3, cor)[2,]
  se <- sd(sample_stats)
  male_normal_CI = rbind(male_normal_CI, cbind(theta_hat - zval*se, theta_hat + zval*se)) 
} 
male_normal_CI$B <- B
male_normal_CI$SampleStatistic <- rep(theta_hat, length(B))
male_normal_CI$Length <- male_normal_CI$V2 - male_normal_CI$V1
male_normal_CI$Shape <- abs(male_normal_CI$V2 - theta_hat)/abs(male_normal_CI$V1 - theta_hat)
male_normal_CI <- male_normal_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(male_normal_CI, row.names = FALSE)
```

When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is symmetric as expected.These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.

**2. Basic Bootstrap CI**

```{r}
set.seed(405568250)

B = c(200, 500, 1000, 2000, 5000, 10000) 
male_basic_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
theta_hat <- cor(male_cats)[2]

for (k in B) {
  boot.samples.male = bootsampling(male_cats, k)
  sample_stats = apply(boot.samples.male, 3, cor)[2,]
  se <- sd(sample_stats)
  male_basic_CI = rbind(male_basic_CI, cbind(2*theta_hat - quantile(sample_stats, probs = (1 - alpha/2)), 2*theta_hat - quantile(sample_stats, probs = (alpha/2)))) 
} 
male_basic_CI$B <- B
male_basic_CI$SampleStatistic <- rep(theta_hat, length(B))
male_basic_CI$Length <- male_basic_CI$V2 - male_basic_CI$V1
male_basic_CI$Shape <- abs(male_basic_CI$V2 - theta_hat)/abs(male_basic_CI$V1 - theta_hat)
male_basic_CI <- male_basic_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(male_basic_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is slightly asymmetric with a larger range of values being between our sample statistic and UB. In general, the length of our male CI seem to be smaller than the female ones. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


**3. Percentile Bootstrap CI**

```{r}
set.seed(405568250)

B = c(200, 500, 1000, 2000, 5000, 10000) 
male_percentile_CI = data.frame(matrix(ncol = 6, nrow = 0))
alpha = 0.05
theta_hat <- cor(male_cats)[2]

for (k in B) {
  boot.samples.male = bootsampling(male_cats, k)
  sample_stats = apply(boot.samples.male, 3, cor)[2,]
  male_percentile_CI = rbind(male_percentile_CI, cbind(quantile(sample_stats, probs = alpha/2), quantile(sample_stats, probs = (1 - (alpha/2))))) 
} 
male_percentile_CI$B <- B
male_percentile_CI$SampleStatistic <- rep(theta_hat, length(B))
male_percentile_CI$Length <- male_percentile_CI$V2 - male_percentile_CI$V1
male_percentile_CI$Shape <- abs(male_percentile_CI$V2 - theta_hat)/abs(male_percentile_CI$V1 - theta_hat)
male_percentile_CI <- male_percentile_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(male_percentile_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is now slightly asymmetric with a larger range of values being between our sample statistic and LB along with a larger degree of asymmetry. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.


**4. Bootstrap-t (Studentized) Bootstrap CI**

```{r}
set.seed(405568250)

B = c(200, 500, 1000, 2500) 
male_studentized_CI = data.frame(matrix(ncol = 4, nrow = 0))
alpha = 0.05
theta_hat <- cor(male_cats)[2]

for (k in B) {
  boot.samples.male <- bootsampling(male_cats, k)
  theta_b = apply(boot.samples.male, 3, cor)[2,]
  boot.samples.SE = rep(0, dim(boot.samples.male)[3])
  se <- sd(theta_b)
  for (l in 1:dim(boot.samples.male)[3]) {
    iterated.samples <- bootsampling(boot.samples.male[,,l], 100)
    theta_l <- apply(iterated.samples, 3, cor)[2,]
    boot.samples.SE[l] <- sd(theta_l)
  }
  t_b <- (theta_b - theta_hat)/boot.samples.SE
  qval = quantile(t_b, probs = c(alpha/2, 1 - alpha/2)) 
  male_studentized_CI = rbind(male_studentized_CI, cbind(theta_hat - qval[2]*se, theta_hat - qval[1]*se))
} 
male_studentized_CI$B <- B
male_studentized_CI$SampleStatistic <- rep(theta_hat, length(B))
male_studentized_CI$Length <- male_studentized_CI$V2 - male_studentized_CI$V1
male_studentized_CI$Shape <- abs(male_studentized_CI$V2 - theta_hat)/abs(male_studentized_CI$V1 - theta_hat)
male_studentized_CI <- male_studentized_CI %>% select(SampleStatistic, B, V1, V2, Length, Shape) %>% rename(Lower = V1) %>% rename(Upper = V2)
print.data.frame(male_studentized_CI, row.names = FALSE)
```


When comparing our CIs to our sample statistic we can see that our CIs are all relatively stable and that an increase in B does not greatly change our intervals. Our shape is now even more asymmetric with a larger range of values being between our sample statistic and LB. These values of B were chosen based on samples from lecture where we are able to showcase a wide variety of bootstrap samples.
